---
title: "Evaluating Transmission Reconstruction Accuracy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluating Transmission Reconstruction Accuracy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(transNetwork)
library(TransPhylo)
```

## Introduction

This vignette demonstrates how to evaluate the accuracy of the **TransPhylo** inference engine using the utility functions in `transNetwork`. We simulate a "ground truth" outbreak, perform Bayesian inference, and then compare the results to see how many infector-infectee pairs were correctly identified.

## The Simulation Loop

In this example, we run `numsim` iterations. For each iteration, we simulate a small outbreak of 10 individuals.

```{r simulation}
set.seed(42)
numsim <- 10
accuracy_scores <- numeric(numsim)

for (i in 1:numsim) {
    # 1. Simulate an outbreak with R0 = 1.0 (infection_rate / removal_rate)
    outbreak <- simulate_outbreak(infection_rate = 3.0, removal_rate = 3.0, target_size = 10)
    numcase <- nrow(outbreak)

    # 2. Build the phylogenetic time tree (the input for TransPhylo)
    timetree <- build_timetree(outbreak, plot = FALSE)
    
    # 3. TransPhylo inference
    ptree <- TransPhylo::ptreeFromPhylo(timetree, dateLastSample = 2007.964)
    res <- TransPhylo::inferTTree(ptree, mcmcIterations = 10000, w.shape = 10, w.scale = 0.1, dateT = 2008)

    # 4. Process Who-Infected-Whom (WIW) matrix
    matrixwiw <- TransPhylo::computeMatWIW(res, burnin = 0.5)
    
    # Use transNetwork utility to convert matrix to edge list
    netedges <- transmission_edges_matrixwiw(matrixwiw)
    
    # 5. Calculate Accuracy
    # Compare the inferred infector (Column 2 of netedges) with the truth (Column 2 of outbreak)
    correct_matches <- length(which(as.numeric(netedges[, 2]) - outbreak[, 2] == 0)) 
    accuracy_scores[i] <- correct_matches / (numcase-1)
}
```

## Results and Analysis

Once the simulations are complete, we can calculate the mean accuracy. This metric represents the probability that the model correctly identifies the true source of infection for a given case.

```{r results}
mean_acc <- mean(accuracy_scores)
print(paste("Average Reconstruction Accuracy:", round(mean_acc, 3)))
```

## Visualizing the Reconstructed Network

Finally, we can use `transplot` to visualize the network from the final simulation. In this plot, the source is highlighted in grey, and edge colors indicate the posterior probability of the transmission link.

```{r visualization, fig.width=7, fig.height=7}
# Plotting the last simulated network
transplot(netedges, style = 1)
```
